Biomass trend maps

### 2019-06-24 analysis/VOCC/biomass-trends.Rmd
sdmTMB and predict working with added covariates

Region of interest
```{r}
region <- "Both odd year surveys"
region <- "West Coast Vancouver Island"
region <- "West Coast Haida Gwaii"
# region <- "Queen Charlotte Sound"
# region <- "Hecate Strait"
# region <- "Both West Coast surveys"
region <- "All synoptic surveys" # Probably not possible
```

Is a new spatial needed?
```{r}
update_spatial_model <- TRUE
# update_spatial_model <- FALSE
priors <- FALSE
# priors <- TRUE
```

@@

  # bath <- readRDS("data/bathymetry-data")
  # dat <- bath$data %>% scale_survey_predictors()

  newcov <- readRDS("data/new_covariates.rds")
  dat <- newcov %>% gfranges::scale_survey_predictors()

@@

  nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>%
    mutate(year = as.numeric(year))
  # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))

  saveRDS(nd_all, file = paste0("data/", spp, "/nd_all_synoptic.rds"))
}
```

@@

Run sdmTMB model (if not done previously).
```{r eval=FALSE}
if (update_spatial_model) {
  biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"))
  covars <- readRDS("event-covariates.rds")
  data <- left_join(biomass, covars)

@@

  if (any(names(data) == "adult_density")) {
    starttime1 <- Sys.time()
    adult_biomass <- sdmTMB::sdmTMB(data,
      adult_density ~ 0 + as.factor(year),
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    saveRDS(adult_biomass, file = paste0("data/", spp, "/model-mature-biomass-", spp, "-", ssid_string, "-priors-", priors, time1, ".rds"))

    starttime2 <- Sys.time()
    imm_biomass <- sdmTMB::sdmTMB(data,
      imm_density ~ 0 + as.factor(year),
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    saveRDS(imm_biomass, file = paste0("data/", spp, "/model-imm-biomass-", spp, "-", ssid_string, "-priors-", priors, time2, ".rds"))
  } else {
    starttime3 <- Sys.time()
    total_biomass <- sdmTMB::sdmTMB(dat,
      density ~ 0 + as.factor(year),
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE, ,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, "-", ssid_string, "-priors-", priors, ".rds"))
  }
}
```

@@

if (!exists("adult_predictions")) {
  if (length(unique(fish$maturity_code)) > 2) {
    nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
    nd <- nd_all %>% filter(ssid %in% model_ssid)
    nd$year <- as.integer(nd$year)

@@

    if (!exists(predictions)) {
      nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
      nd <- nd_all %>% filter(ssid %in% model_ssid)
      nd$year <- as.integer(nd$year)

>>>>

Is a new spatial needed?
```{r}
update_spatial_model <- TRUE
priors <- FALSE
# priors <- TRUE

# Choose region to update model for:

region <- "Both odd year surveys"
region <- "West Coast Vancouver Island"
region <- "West Coast Haida Gwaii"

# region <- "Queen Charlotte Sound"
# region <- "Hecate Strait"
# region <- "Both West Coast surveys"
```

If not...
```{r}
update_spatial_model <- FALSE
priors <- FALSE
# priors <- TRUE
region <- "All synoptic surveys" 
```

@@

  bath <- readRDS("data/bathymetry-data")
  dat <- bath$data %>% scale_survey_predictors()

@@

  nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>% 
    mutate(year = as.numeric(year))

  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky)
  nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))

  # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))

  saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
}
```

@@

Run sdmTMB model (if not done previously).
```{r eval=FALSE}
if (update_spatial_model) {
  biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass,covars)

@@

  if (any(names(data) == "adult_density")) {
    starttime1 <- Sys.time()
    adult_biomass <- sdmTMB::sdmTMB(data,
      adult_density ~ 0 + as.factor(year) + trawled + muddy + rocky,
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    saveRDS(adult_biomass, file = paste0("data/", spp, "/model-mature-biomass-", spp, "-", ssid_string, "-priors-", priors, time1, "-newcovs.rds"))

    starttime2 <- Sys.time()
    imm_biomass <- sdmTMB::sdmTMB(data,
      imm_density ~ 0 + as.factor(year) + trawled + muddy + rocky,
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    saveRDS(imm_biomass, file = paste0("data/", spp, "/model-imm-biomass-", spp, "-", ssid_string, "-priors-", priors, time2, "-newcovs.rds"))
  } else {
    starttime3 <- Sys.time()
    total_biomass <- sdmTMB::sdmTMB(dat,
      density ~ 0 + as.factor(year)+ trawled + muddy + rocky,
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE, ,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, "-", ssid_string, "-priors-", priors, "-newcovs.rds"))
  }
}
```

@@

if (!exists("adult_predictions")) {
  if (length(unique(fish$maturity_code)) > 2) {
    nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
    nd <- nd_all %>% filter(ssid %in% model_ssid) 
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)

@@

    if (!exists(predictions)) {
      nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
      nd <- nd_all %>% filter(ssid %in% model_ssid)
      nd <- na.omit(nd)
      nd$year <- as.integer(nd$year)
